Software architectures to integrate workflow engines in science gateways

### 2016-10-12 answers.tex
Finished new section 6

\usepackage{hyperref}
\newenvironment{review}%          environment name
{\textbf{Reviewer comment:}\begin{quote}}% begin code
{\end{quote}}%  
\newcommand{\answer}[1]{\textbf{Answer:} #1}
\newcommand{\todo}[1]{\color{red}#1\color{black}}

@@

\answer{We thank the reviewer for the thorough review and useful comments.

@@

  We added the following paragraph to the Introduction to reflect this discussion:}

@@

@@

\answer{The actual workflow engine deployed in the infrastructure
  requires the gateway to directly control its own sub-tasks. We
  revised the wording of Section 2.4 to clarify this point.}

@@

\answer{S1 denotes the ability of the system to start multiple engine
  instances to execute different workflows, while S2 denotes the
  possibility to execute a single workflow with multiple engines. We
  revised the wording in the description of S1 (page 11) to clarify
  this point. We find the current description of S2 clear enough on
  this point: ``[S2] measures the ability to distribute the execution of
  \emph{a single workflow} among different engine instances.''}

@@

\answer{We introduced E3 precisely to account for the additional
  complexity introduced by interaction (g) when adding new
  workflows.

@@

\revised{Workflow conversion is presented here as a pattern to
integrate workflow \emph{engines} in a science gateway, through
workflow format conversion from a native format to the science gateway
format.}
}

@@

This sentence complements the note already made in section 3.7:
``Note, however,
that implementing interaction g can require sub-
stantial work depending on the complexity of the
workflow language used by the new engine.``

@@

\answer{As discussed previously, we share the reviewer's concerns
  about the feasibility of workflow conversion (see comments on
  interaction (g) in Section 5.2). The performance of workflow
  conversion is strongly related to the availability of a validated
  interaction (g), which is very dependent on the particular workflow
  engines under consideration. We believe that the issue raised by the
  reviewer basically comes down to the weight given to software
  components and interactions in the analysis. However, using
  different weights for different components or interactions would
  open an infinite number of configurations for discussion. For
  instance, in some cases, the workflow engine already comes with a
  workflow service which therefore doesn't have to be included. To
  address this issue, we shared a spreadsheet at
  \url{https://frama.link/LrPRsCQp} that allows editing the following
  weights:
  \begin{itemize}
  \item Weight given to software components and interactions in
  ``integration'', ``extensibility'' and ``robustness'' criteria.
  \item Weight given to metrics in a criterion (e.g., E$_1$ vs. E$_2$ in
  ``extensibility'').
  \item Weight given to criteria in overall evaluation of Table 3.
\end{itemize}
The paper references this spreadsheet wherever equal weights are
assumed. The readers and reviewers are invited to rely on this
spreadsheet for any issue related to the importance given to a
specific component, interaction, metric or criterion in the
analysis. In the paper, we assumed that all weights equal 1, which we
believe is a reasonable but limited choice.
}

@@

\answer{The reviewer is right that re-purposing software components
  and interactions can lead to revisit the evaluation. We updated the following sentence in section 5.2. to better acknowledge this point:}

@@

\revised{However, when a new type of workflow engine is added, it is always possible to reuse either an existing available workflow service or interaction \texttt{d} (when a new workflow service has to be developed). Therefore, \texttt{E$_1$}=3}.

@@

\answer{While we agree with the reviewer that re-usability should
  definitely be taken into account in E$_1$, we do not see how it
  could help S$_1$ and S$_2$.

@@

  For S$_2$, we cannot think of a practical reason why the service
  invocation model would help distribute a single workflow execution
  in multiple engine instances. In our opinion, this feature is
  exclusive to the nested workflow model. }

@@

We hope that the update made in the extensibility scores and the
explanations given for scalability convinced the reviewer.

@@

\answer{We updated the definition of R$_1$ and R$_2$ to include
  positive examples as well as negative ones. In addition, the
  specific components included in these metrics are represented in red
  on Figure 2.}

@@

\texttt{c}, \texttt{c$_1$} and \texttt{c$_2$} are three different
interactions, which explains the confusion. \texttt{c$_2$} is always
specific to workflow execution because it is defined as the
interaction between the workflow engine and the infrastructure, used
to transfer workflow data (inputs, outputs or intermediate
results). \texttt{c$_1$} is not specific to workflow execution because
it is used by the science gateway to transfer any data regardless of
workflow executions. Letter \texttt{c} was used on the task
encapsulation diagram in Figure 2 to represent that the science gateway transfers
both user and workflow data. We now use \texttt{c$_1$} for all the
data transfers initiated by the science gateway, and \texttt{c$_2$}
for all the data transfers initiated by the workflow engine,
regardless of the type of data transferred. Consequently, we now use
\texttt{c$_1$} instead of \texttt{c} on the task encapsulation diagram in Figure 2.

@@

Specific components (\texttt{R$_1$}): number of specific components involved in workflow executions. \revised{For instance, the
  workflow service is a specific component because its main objective
  is to execute workflows. On the contrary, the science gateway is
  \emph{not} specific to workflow execution since it is used for a
  variety of other functions such as user authentication, data
  transfer, etc.}

Specific interactions (\texttt{R$_2$}): number of specific
  interactions involved in workflow executions. \revised{For
    instance, the pool-agent interaction (interaction \texttt{f}) is
    specific to workflow execution while data control (interaction
    \texttt{c$_1$}) is \emph{not} because it is used by the science gateway to transfer
    user data regardless of a particular workflow execution}.

@@

\answer{No it doesn't. We hope that the new examples in the definition of
R$_1$ and R$_2$ and the removal of any reference to interaction
\texttt{c} clarifies this point.}

@@

\answer{Yes, as for any other metric, high values indicate worse
performance. This is explained at the beginning of section 3:
``Criteria break down into specific metrics where lower value
indicates better performance.''}

@@

\begin{review}
To conclude, the general feeling is that the definition of the metrics and the interaction patterns are not rigorous enough to provide an interpretation of the final evaluation which is entirely useful for the reader. Probably, some more information on how "Future Generation" systems should be designed, starting from the general overview provided, with more technical understanding of the system complexities, could represent a more valuable contribution to the main aim of this particular publication.
\end{review}

@@

\answer{This was done as suggested.}

@@

\answer{The task submission framework is part of the infrastructure
  and its nature doesn't influence the architecture used to integrate
  workflow engines in the science gateway. Formally, it is part of
  interaction \texttt{b} (see definition in 2.1).

@@

  We found this scenario interesting and we evaluated it in our
  framework. The analysis is reported in the new section 6 of the
  paper.}

@@

\answer{complexity is used as a measure of reliability}

>>>>

\usepackage{hyperref}
\newenvironment{review}%          environment name
{\textbf{Reviewer comment:}\begin{quote}}% begin code
{\end{quote}}%  
\newenvironment{answer}%          environment name
{\textbf{Answer:}\begin{small}\begin{quote}}% begin code
{\end{quote}\end{small}}%

\newcommand{\todo}[1]{\color{red}#1\color{black}}

@@

\begin{answer}

  We thank the reviewer for the thorough review and useful comments.

@@

  We added the following paragraph to the Introduction to reflect this discussion:

@@

\end{answer}

@@

\begin{answer}
  The actual workflow engine deployed in the infrastructure
  requires the gateway to directly control its own sub-tasks. We
  revised the wording of Section 2.4 to clarify this point.
\end{answer}

@@

\begin{answer}
  S1 denotes the ability of the system to start multiple engine
  instances to execute different workflows, while S2 denotes the
  possibility to execute a single workflow with multiple engines. We
  revised the wording in the description of S1 (page 11) to clarify
  this point. We find the current description of S2 clear enough on
  this point: ``[S2] measures the ability to distribute the execution of
  \emph{a single workflow} among different engine instances.''
\end{answer}

@@

\begin{answer}
  We introduced E3 precisely to account for the additional
  complexity introduced by interaction (g) when adding new
  workflows.

@@

  \revised{Workflow conversion is presented here as a pattern to
integrate workflow \emph{engines} in a science gateway, through
workflow format conversion from a native format to the science gateway
format.}

@@

This sentence complements the note already made in section 3.7:
``Note, however,
that implementing interaction g can require sub-
stantial work depending on the complexity of the
workflow language used by the new engine.``
\end{answer}

@@

\begin{answer}

  As discussed previously, we share the reviewer's concerns
  about the feasibility of workflow conversion (see comments on
  interaction (g) in Section 5.2). The performance of workflow
  conversion is strongly related to the availability of a validated
  interaction (g), which is very dependent on the particular workflow
  engines under consideration. We believe that the issue raised by the
  reviewer basically comes down to the weight given to software
  components and interactions in the analysis. However, using
  different weights for different components or interactions would
  open an infinite number of configurations for discussion. For
  instance, in some cases, the workflow engine already comes with a
  workflow service which therefore doesn't have to be included. To
  address this issue, we shared a spreadsheet at
  \url{https://frama.link/LrPRsCQp} that allows editing the following
  weights:
  \begin{itemize}
  \item Weight given to software components and interactions in
  ``integration'', ``extensibility'' and ``robustness'' criteria.
  \item Weight given to metrics in a criterion (e.g., E$_1$ vs. E$_2$ in
  ``extensibility'').
  \item Weight given to criteria in overall evaluation of Table 3.
\end{itemize}
The paper references this spreadsheet wherever equal weights are
assumed. The readers and reviewers are invited to rely on this
spreadsheet for any issue related to the importance given to a
specific component, interaction, metric or criterion in the
analysis. In the paper, we assumed that all weights equal 1, which we
believe is a reasonable but limited choice.

\end{answer}

@@

\begin{answer}

The reviewer is right that re-purposing software components
  and interactions can lead to revisit the evaluation. We updated the following sentence in section 5.2. to better acknowledge this point:

@@

\revised{However, when a new type of workflow engine is added, it is always possible to reuse either an existing available workflow service or interaction \texttt{d} (when a new workflow service has to be developed). Therefore, \texttt{E$_1$}=3}.
\end{answer}

@@

\begin{answer}
  While we agree with the reviewer that re-usability should
  definitely be taken into account in E$_1$, we do not see how it
  could help S$_1$ and S$_2$.

@@

  For S$_2$, we cannot think of a practical reason why the service
  invocation model would help distribute a single workflow execution
  in multiple engine instances. In our opinion, this feature is
  exclusive to the nested workflow model.
\end{answer}

@@

\begin{answer}
We hope that the update made in the extensibility scores and the
explanations given for scalability convinced the reviewer.
\end{answer}

@@

\begin{answer}
We updated the definition of R$_1$ and R$_2$ to include
  positive examples as well as negative ones. In addition, the
  specific components included in these metrics are represented in red
  on Figure 2.
\end{answer}

@@

\begin{answer}
\texttt{c}, \texttt{c$_1$} and \texttt{c$_2$} are three different
interactions, which explains the confusion. \texttt{c$_2$} is always
specific to workflow execution because it is defined as the
interaction between the workflow engine and the infrastructure, used
to transfer workflow data (inputs, outputs or intermediate
results). \texttt{c$_1$} is not specific to workflow execution because
it is used by the science gateway to transfer any data regardless of
workflow executions. Letter \texttt{c} was used on the task
encapsulation diagram in Figure 2 to represent that the science gateway transfers
both user and workflow data. We now use \texttt{c$_1$} for all the
data transfers initiated by the science gateway, and \texttt{c$_2$}
for all the data transfers initiated by the workflow engine,
regardless of the type of data transferred. Consequently, we now use
\texttt{c$_1$} instead of \texttt{c} on the task encapsulation diagram in Figure 2.

@@

\begin{itemize}
\item Specific components (\texttt{R$_1$}): number of specific components involved in workflow executions. \revised{For instance, the
  workflow service is a specific component because its main objective
  is to execute workflows. On the contrary, the science gateway is
  \emph{not} specific to workflow execution since it is used for a
  variety of other functions such as user authentication, data
  transfer, etc.}
\item Specific interactions (\texttt{R$_2$}): number of specific
  interactions involved in workflow executions. \revised{For
    instance, the pool-agent interaction (interaction \texttt{f}) is
    specific to workflow execution while data control (interaction
    \texttt{c$_1$}) is \emph{not} because it is used by the science gateway to transfer
    user data regardless of a particular workflow execution}.
\end{itemize}
\end{answer}

@@

\begin{answer}
No it doesn't. We hope that the new examples in the definition of
R$_1$ and R$_2$ and the removal of any reference to interaction
\texttt{c} clarifies this point.
\end{answer}

@@

\begin{answer}
Yes, as for any other metric, high values indicate worse
performance. This is explained at the beginning of section 3:
``Criteria break down into specific metrics where lower value
indicates better performance.''
\end{answer}

@@

\begin{answer}
  We call ``white-box'' workflow a workflow
  where detailed information about individual tasks is available,
  for instance task statuses, execution logs and resource
  consumption. On the contrary, a ``black-box'' workflow only
  displays global information such as the workflow status and the
  logs of the workflow engine. Fine-grained information about
  workflow tasks helps workflow users (scientists)
  identify errors, it allows gateway administrators to troubleshoot
  executions, and it facilitates debugging by workflow
  developers.

  Multiple software layers affect the accuracy of the information
  reported about the tasks' internals because each layer potentially
  shields or modifies information. Task ids, for instance, may not
  be properly linked to workflow activities if delivered through a
  software layer that hasn't been instrumented for that. \todo{refer
    to olabarriaga2014}.

  The paragraph on fine-grained debugging has been edited accordingly, at the end of Section 3.
\end{answer}

\begin{review}
To conclude, the general feeling is that the definition of the metrics and the interaction patterns are not rigorous enough to provide an interpretation of the final evaluation which is entirely useful for the reader.
\end{review}

\begin{answer}
We hope that the explanations provided in this letter and the updated text in the paper demonstrate that the metrics and definitions are rigorous.
\end{answer}

\begin{review}
 Probably, some more information on how "Future Generation" systems should be designed, starting from the general overview provided, with more technical understanding of the system complexities, could represent a more valuable contribution to the main aim of this particular publication.
\end{review}

\begin{answer}
  We added a section in the paper (current section 6) to show
  how our evaluation framework could be used in the design of new
  systems.
\end{answer}

@@

\begin{answer}
This was done as suggested.
\end{answer}

@@

\begin{answer}
  The task submission framework is part of the infrastructure
  and its nature doesn't influence the architecture used to integrate
  workflow engines in the science gateway. Formally, it is part of
  interaction \texttt{b} (see definition in 2.1).

@@

  We found this scenario interesting and we evaluated it in our
  framework. The analysis is reported in the new section 6 of the
  paper.
\end{answer}

@@

\todo{complexity is used as a measure of reliability}
